Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-07-23 paper_jasa.tex
definition p-value

The paper is organized as follows. Section \ref{sec:visual_test} defines terms as used in visual inference, and describes how to estimate the important quantities from experimental data. The effect of the lineup size and number of observers on the power of the test is discussed in Section \ref{sec:size}. Section \ref{sec:regression} focuses on the application of  visual inference to linear models.  \blue{Section \ref{sec:simulation} describes the three simulation experiments conducted to compare the power of the lineup protocol with the equivalent conventional test and  Section \ref{sec:results} presents the results of the experiments with analysis of the resulting data.} The last section describes the next steps in this research.

@@

\begin{dfn}\label{dfn:pvalue}
The $p$-value of a lineup of size $m$ is  estimated as %given as

@@

\noindent An important difference between conventional and visual testing is that lineups will depend on observers evaluation. Thus, the number of observers affects the estimation of power, where $X$ is the number of observers (out of $K$) who identify the actual data plot from the lineup, and power is estimated by
\[
\widehat{Power}_{V} (\theta) = {Power}_{V, K} (\theta) = 1 - F_{X, \theta} (x_{\alpha} - 1),
\] 
Here $F_{X, \theta}$ is the distribution of $X$, and this depends on which hypothesis is true.
Under the null hypothesis, $X \sim Binom_{K, 1/m}$, leading to:
\[
\widehat{Power}_V(\theta, K)= 1 - Binom_{K, 1/m} (x_\alpha - 1),
\]
where $x_\alpha$ is such that $P(X \ge x_{\alpha}) < \alpha$. 
If the alternative hypothesis is true, with a fixed parameter value $\theta$, we can assume that an individual's probability to identify the data plot depends on the parameter value, and $X_i \sim Binom_{1, p_i(\theta)}$. Assessing an individual's skill to identify the actual data plot will require that an individual evaluates multiple lineups.

>>>>

The paper is organized as follows. Section \ref{sec:visual_test} defines terms as used in visual inference, and describes how to estimate the important quantities from experimental data. The effect of the lineup size and number of observers on the power of the test is discussed in Section \ref{sec:size}. Section \ref{sec:regression} focuses on the application of  visual inference to linear models.  Section \ref{sec:simulation} describes  three  user studies based on simulation experiments conducted to compare the power of the lineup protocol with the equivalent conventional test and  Section \ref{sec:results} presents an analysis of the resulting data. %The last section describes the next steps in this research.

@@

\begin{dfn}\label{dfn:pvalue}
The $p$-value of a lineup of size $m$ evaluated by $K$ observers is  given as

@@

\noindent An important difference between conventional and visual testing is that lineups will depend on observers evaluation. Thus, the number of observers affects the estimation of power, where $X$ is the number of observers (out of $K$) who identify the actual data plot from the lineup, and power is estimated by
\[
\widehat{Power}_{V} (\theta) = {Power}_{V, K} (\theta) = 1 - F_{X, \theta} (x_{\alpha} - 1),
\] 
Here $F_{X, \theta}$ is the distribution of $X$ and $x_\alpha$ is such that $P(X \ge x_{\alpha}) \le \alpha$. Note that the distribution  $F_X$ depends on which hypothesis is true:
under the null hypothesis, $X \sim Binom_{K, 1/m}$, leading to:
\[
Power_V(\theta, K)= 1 - Binom_{K, 1/m} (x_\alpha - 1),
\]
%where $x_\alpha$ is such that $P(X \ge x_{\alpha}) \le \alpha$. 
If the alternative hypothesis is true, with a fixed parameter value $\theta$, we can assume that an individual's probability to identify the data plot depends on the parameter value, and $X_i \sim Binom_{1, p_i(\theta)}$. Assessing an individual's skill to identify the actual data plot will require that an individual evaluates multiple lineups.
